Plant functional type effects on biomass partitioning

### 2014-12-10 manuscript.Rnw

>>>>

\section*{First paragraph - Daniel}

\textbf{Leaf traits drive differences in biomass partitioning among major plant functional types}

The partitioning of plant primary productivity among fleshy leaves and woody stems is a key factor driving carbon cycles and the flow of energy within terrestrial ecosystems. Although researchers have long assumed key plant functional types (PFTs) to differ in biomass partitioning, empirical data have been lacking. Divergent approaches to modelling allocation is thus a major source of uncertainty in terrestrial ecosystem models (TEMs) and global carbon cycles. Using a new database on Biomass partitioning combining data from over 175 studies, we here show a strong effect of PFT on biomass partitioning. Contrary to prior expectation, we show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that such differences relate to systematic shifts in leaf-mass per area, a leading trait within the leaf economics spectrum. Another core trait, leaf area per sapwood area, explained further variation among species in biomass partitioning, but did not differ systematically among PFTs. Combined, these results uncover fundamental differences in how plants are constructed and allow for systematic testing of allocation routines with current TEMs.
